---
title: "Compiled HUD SFMR for Washington State"
author: "Lamar Foster and Delaney Glass"
date: "7/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rio)
library(dplyr)
library(janitor)
library(tidyr)
library(stringr)
```

## Read in the data 

```{r}

# read in data from HUD Small Area Fair Market Rents which can be found here: https://www.huduser.gov/portal/datasets/fmr/smallarea/index.html#2014

hud10_11 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2011f/fy2011_equiv_zc_rents_acs.xls")
hud12 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2012f/small_area_fmrs_fy2012.xls")
hud13 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2013p/small_area_fmrs_fy2013.xls")
hud14 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2014f/small_area_fmrs_fy2014.xls")
hud15 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2015f/small_area_fmrs_fy2015f.xls")
hud16 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2016f/final_fy2016_hypothetical_safmrs.xlsx")
hud17 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2017/FY2017_hypothetical_safmrs.xlsx")
```

## Data from 2012-2014 have matching col names, so for now we will use clean_names for the other dataframes

```{r}
hud10_11 <- hud10_11 %>% clean_names() # use clean_names to make all names lower-case
hud15 <- hud15 %>% clean_names()
hud16 <- hud16 %>% clean_names()
hud17 <- hud17 %>% clean_names()

```

# Combine data from 2012-2014 

```{r}

# create 'year' variable to make binding easier later

hud12 <- hud12 %>% 
  mutate(year = 2012)

hud13 <- hud13 %>% 
  mutate(year = 2013)

hud14 <- hud14 %>% 
  mutate(year = 2014)

# combine 2012 - 2014 data and get rid of extraneous variables
hud12_14 <- bind_rows(hud12, hud13, hud14) %>% clean_names() %>% 
  select(-c(state, state_name, county))

```

## make 'year' variables for remaining df's, rename for convergence, and drop extraneous variables 

```{r}

hud10_11 <- hud10_11 %>% 
  select(-c(state, county)) %>% 
rename(
cbsa = cbsamet) %>% 
  mutate(year = 2011)
```

```{r}
hud15 <- hud15 %>% 
  select(-c(state, statename, county)) %>% 
rename(
cbsa = cbsamet,
county_name = cntyname,
zip = zipcode,
cbsa_name = cbnsmcnm) %>% 
  mutate(year = 2015)
```

```{r}
hud16 <- hud16 %>% 
  select(-c(fips_state_code, fips_county_code, statename)) %>% 
rename(
cbsa = metro_code,
cbsa_name = metro_name,
zip = zip_code) %>% 
  mutate(year = 2016)
```

```{r}
hud17 <- hud17 %>% 
rename(
cbsa = metro_code,
cbsa_name = metro_name,
zip = zip_code) %>% 
  mutate(year = 2017)

```

```{r}

#Add county to hud_17 data frame 
hud17 <- left_join(hud17, hud16, by = "zip") %>% 
  select(ends_with("x"), county_name, zip) %>% 
  rename_at(.vars = vars(ends_with(".x")), .funs = funs(sub("[.]x$", "",.)))
```

## combine all dataframes together so that it shows all years rental prices (studio-four bedrooms) across all years

```{r}
hud_full <- plyr::rbind.fill(hud12_14, hud10_11, hud15, hud16, hud17)
```

## filter out only Washington State data 

```{r}
# Filter data for only Washington State 
hud_WA <- hud_full %>% 
  separate(cbsa_name, into = c("city", "region"), sep = ",") %>%
  mutate(region = str_trim(region)) %>% 
  filter(region == "WA MSA") %>% 
  rename(
    region = city,
    state = region) %>%
  mutate(state = "WA")
```

```{r}
# make long format

#Wrangle data to make it long
hud_long <- hud_WA %>% 
  rename(
    county = county_name,
    Studio = area_rent_br0,
    `One Bedroom` = area_rent_br1,
    `Two Bedrooms` = area_rent_br2,
    `Three Bedrooms` = area_rent_br3,
    `Four Bedrooms` = area_rent_br4) %>% 
  pivot_longer(c(Studio, `One Bedroom`, `Two Bedrooms`, `Three Bedrooms`, `Four Bedrooms`), names_to = "Bedrooms", values_to = "Rent")

```

```{r}

# Make region and bedroom a factor with specific levels 
hud_long$Bedrooms <- factor(hud_long$Bedrooms, 
                         levels = c("Studio", "One Bedroom", "Two Bedrooms", "Three Bedrooms", "Four Bedrooms"))
hud_long$region <- factor(hud_long$region)
hud_long$year <- as.integer(hud_long$year)
hud_long <- hud_long %>% mutate(zip = parse_number(zip))
hud_long[hud_long == "Kennewick-Richland" ] <-  "Kennewick-Pasco-Richland"
hud_long[hud_long == "Olympia" ] <-  "Olympia-Tumwater"
hud_long[hud_long == "Spokane" ] <-  "Spokane-Spokane Valley"
hud_long[hud_long == "Wenatchee" ] <-  "Wenatchee-East Wenatchee"

```

## Write .csv file 
```{r}

write.csv(hud_long, "~/Dropbox/UW 2020-2021/DSSG 2021/housing-wage/hud_sfmr_final_WA.csv")

```