---
title: "Compiled HUD SFMR for Washington State"
author: "Lamar Foster and Delaney Glass"
date: "7/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rio)
library(dplyr)
library(janitor)
library(tidyr)
library(stringr)
```

## Read in the data 

```{r}

# read in data from HUD Small Area Fair Market Rents which can be found here: https://www.huduser.gov/portal/datasets/fmr/smallarea/index.html#2014

hud10_11 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2011f/fy2011_equiv_zc_rents_acs.xls")
hud12 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2012f/small_area_fmrs_fy2012.xls")
hud13 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2013p/small_area_fmrs_fy2013.xls")
hud14 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2014f/small_area_fmrs_fy2014.xls")
hud15 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2015f/small_area_fmrs_fy2015f.xls")
hud16 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2016f/final_fy2016_hypothetical_safmrs.xlsx")
hud17 <- import(file = "https://www.huduser.gov/portal/datasets/fmr/fmr2017/FY2017_hypothetical_safmrs.xlsx")
```

## Data from 2012-2014 have matching col names, so for now we will use clean_names for the other dataframes

```{r}
hud10_11 <- hud10_11 %>% clean_names() # use clean_names to make all names lower-case
hud15 <- hud15 %>% clean_names()
hud16 <- hud16 %>% clean_names()
hud17 <- hud17 %>% clean_names()

```

# Combine data from 2012-2014 

```{r}

# create 'year' variable to make binding easier later

hud12 <- hud12 %>% 
  mutate(year = 2012)

hud13 <- hud13 %>% 
  mutate(year = 2013)

hud14 <- hud14 %>% 
  mutate(year = 2014)

# combine 2012 - 2014 data and get rid of extraneous variables
hud12_14 <- bind_rows(hud12, hud13, hud14) %>% clean_names() %>% 
  select(-c(state, state_name, county))

```

## make 'year' variables for remaining df's, rename for convergence, and drop extraneous variables 

```{r}

hud10_11 <- hud10_11 %>% 
  select(-c(state, county)) %>% 
rename(
cbsa = cbsamet) %>% 
  mutate(year = 2011)
```

```{r}
hud15 <- hud15 %>% 
  select(-c(state, statename, county)) %>% 
rename(
cbsa = cbsamet,
county_name = cntyname,
zip = zipcode,
cbsa_name = cbnsmcnm) %>% 
  mutate(year = 2015)
```

```{r}
hud16 <- hud16 %>% 
  select(-c(fips_state_code, fips_county_code, statename)) %>% 
rename(
cbsa = metro_code,
cbsa_name = metro_name,
zip = zip_code) %>% 
  mutate(year = 2016)
```

```{r}
hud17 <- hud17 %>% 
rename(
cbsa = metro_code,
cbsa_name = metro_name,
zip = zip_code) %>% 
  mutate(year = 2017)
```

## combine all dataframes together so that it shows all years rental prices (studio-four bedrooms) across all years

```{r}
hud_full <- plyr::rbind.fill(hud12_14, hud10_11, hud15, hud16, hud17)
```

## filter out only Washington State data 

```{r}
hud_full %>% 
  separate(cbsa_name, into = c("city", "state"), sep = ",") %>%
  mutate(state = str_trim(state)) %>% 
  filter(state == "WA MSA") %>% 
  View()
```

## Write .csv file 
```{r}

write.csv(hud_full, "~/Dropbox/UW 2020-2021/DSSG 2021/housing-wage/hud_sfmr_2010_2017.csv")
```